Appendix 2: Simulating 20-species landscapes

### 2016-03-14 Harris_2015/show-results.tex
final before resubmission to Ecology?

\section{A: Import packages and data}\label{a-import-packages-and-data}

@@

@@

\subsection{\texorpdfstring{Plot estimates versus ``true'' values across
methods and lanscape
sizes}{Plot estimates versus true values across methods and lanscape sizes}}\label{plot-estimates-versus-true-values-across-methods-and-lanscape-sizes}

The gray line indicates the 1:1 line, where models recover an unbiased
estimate of the ``true'' \(\beta\) coefficients. The GLM and Markov
network methods estimate the coefficients on the original scale, and
approach unbiasedness as the number of sites increases. The four methods
based on (partial) correlations are constrained to return values between
-1 and 1, and do not return values on the original scale. The covariance
and partial covariance (which are not scaled in this way) may avoid this
problem in future analyses (Loh and Wainwright 2013; see main text for
full citation).

Note that the null model (Pairs) is not included in these graphs, as its
test statistic depends strongly on the size of the landscape and fills a
range that would not fit on these graphs.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{filter}\NormalTok{(x, simulation_type ==}\StringTok{ "no_env"}\NormalTok{, method !=}\StringTok{ "null"}\NormalTok{),}
       \KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =} \NormalTok{truth, }\DataTypeTok{y =} \NormalTok{estimate, }\DataTypeTok{color =} \KeywordTok{factor}\NormalTok{(n_sites))) +}\StringTok{ }
\StringTok{  }\KeywordTok{facet_wrap}\NormalTok{(~method) +}
\StringTok{  }\KeywordTok{geom_point}\NormalTok{(}\DataTypeTok{alpha =} \FloatTok{0.01}\NormalTok{) +}
\StringTok{  }\KeywordTok{geom_abline}\NormalTok{(}\DataTypeTok{intercept =} \DecValTok{0}\NormalTok{, }\DataTypeTok{slope =} \DecValTok{1}\NormalTok{, }\DataTypeTok{size =} \DecValTok{2}\NormalTok{, }\DataTypeTok{col =} \StringTok{"darkgray"}\NormalTok{) +}\StringTok{ }
\StringTok{  }\KeywordTok{geom_smooth}\NormalTok{(}\DataTypeTok{method =} \StringTok{"lm"}\NormalTok{, }\DataTypeTok{fullrange =} \OtherTok{FALSE}\NormalTok{) +}\StringTok{ }
\StringTok{  }\KeywordTok{coord_equal}\NormalTok{() +}\StringTok{ }
\StringTok{  }\KeywordTok{scale_color_discrete}\NormalTok{(}\DataTypeTok{name =} \StringTok{"n_sites"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{show-results_files/figure-latex/unnamed-chunk-8-1.png}

@@

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{pairs_summary =}\StringTok{ }\NormalTok{x[x$method ==}\StringTok{ "null"} \NormalTok{&}\StringTok{ }\NormalTok{!}\KeywordTok{grepl}\NormalTok{(}\StringTok{"pop"}\NormalTok{, x$rep_name), ]}
\NormalTok{markov_summary =}\StringTok{ }\NormalTok{x[x$method ==}\StringTok{ "Markov network"} \NormalTok{&}\StringTok{ }\NormalTok{!}\KeywordTok{grepl}\NormalTok{(}\StringTok{"pop"}\NormalTok{, x$rep_name), ]}

@@

The \texttt{error\_smoother} is a function that fits a generalized
additive model to estimate the probability that a model estimate will
match some criterion (e.g. statistical significance) for some parameter
as a function of that parameter's ``true'' value. This function is used
in Figure 4C and in calculating Type I error rates below.

@@

\NormalTok{error_smoother =}\StringTok{ }\NormalTok{function(data, f, }\DataTypeTok{values =} \NormalTok{truth_seq)\{}
  \KeywordTok{predict}\NormalTok{(}
    \KeywordTok{gam}\NormalTok{(}
      \KeywordTok{f}\NormalTok{(data) ~}\StringTok{ }\KeywordTok{s}\NormalTok{(truth),}
      \DataTypeTok{data =} \NormalTok{data,}
      \DataTypeTok{family =} \NormalTok{binomial}
    \NormalTok{),}
    \KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{truth =} \NormalTok{values),}
    \DataTypeTok{type =} \StringTok{"response"}
  \NormalTok{)}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

@@

\CommentTok{# Calculate p(confidently wrong) for each model as a function of}
\CommentTok{# the true values}
\NormalTok{confidently_wrong =}\StringTok{ }\NormalTok{function(data)\{}
  \NormalTok{(data$sig_pos &}\StringTok{ }\NormalTok{data$truth <}\StringTok{ }\DecValTok{0}\NormalTok{) |}\StringTok{ }\NormalTok{(data$sig_neg &}\StringTok{ }\NormalTok{data$truth >}\StringTok{ }\DecValTok{0}\NormalTok{)}
\NormalTok{\}}
\NormalTok{y_markov =}\StringTok{ }\KeywordTok{error_smoother}\NormalTok{(markov_summary, confidently_wrong)}
\NormalTok{y_pairs =}\StringTok{ }\KeywordTok{error_smoother}\NormalTok{(pairs_summary, confidently_wrong)}

@@

\KeywordTok{plot}\NormalTok{(}
  \NormalTok{truth_seq,}
  \NormalTok{y_pairs,}
  \DataTypeTok{type =} \StringTok{"l"}\NormalTok{,}
  \DataTypeTok{xlab =} \StringTok{"}\CharTok{\textbackslash{}"}\StringTok{True}\CharTok{\textbackslash{}"}\StringTok{ interaction strength"}\NormalTok{,}
  \DataTypeTok{ylab =} \StringTok{"P(confidently predict wrong sign)"}\NormalTok{,}
  \DataTypeTok{bty =} \StringTok{"l"}\NormalTok{,}
  \DataTypeTok{yaxs =} \StringTok{"i"}\NormalTok{,}
  \DataTypeTok{col =} \DecValTok{2}\NormalTok{,}
  \DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, .}\DecValTok{4}\NormalTok{),}
  \DataTypeTok{lwd =} \DecValTok{2}
\NormalTok{)}
\KeywordTok{mtext}\NormalTok{(}\StringTok{"C. Error rate vs.}\CharTok{\textbackslash{}n}\StringTok{interaction strength"}\NormalTok{, }\DataTypeTok{side =} \DecValTok{3}\NormalTok{, }\DataTypeTok{adj =} \DecValTok{0}\NormalTok{, }
      \DataTypeTok{font =} \DecValTok{2}\NormalTok{, }\DataTypeTok{line =} \FloatTok{1.2}\NormalTok{, }\DataTypeTok{cex =} \NormalTok{.}\DecValTok{9}\NormalTok{) }
\KeywordTok{lines}\NormalTok{(truth_seq, y_markov, }\DataTypeTok{lwd =} \DecValTok{2}\NormalTok{)}
\KeywordTok{legend}\NormalTok{(}\StringTok{"topleft"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{2}\NormalTok{, }\DataTypeTok{legend =} \KeywordTok{c}\NormalTok{(}\StringTok{"Null model"}\NormalTok{, }\StringTok{"Markov network"}\NormalTok{), }
       \DataTypeTok{col =} \KeywordTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{, }\DecValTok{1}\NormalTok{), }\DataTypeTok{bty =} \StringTok{"n"}\NormalTok{)}
\KeywordTok{dev.off}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

@@

\begin{verbatim}
## [1] 0.1533758
\end{verbatim}

@@

\begin{verbatim}
## [1] 0.02861541
\end{verbatim}

@@

\begin{verbatim}
## [1] 0.4520535
\end{verbatim}

@@

\begin{verbatim}
## [1] 0.2116561
\end{verbatim}

@@

Type I error rates (probability of rejecting the null hypothesis given a
``true'' interaction strength of zero, according to the spline model
defined by \texttt{error\_smoother} above). The smoother focuses on the
64\% of the data points whose ``true'' values fall between -1 and 1,
because its estimates near 0 are more accurate when they aren't affected
by distant data points.

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{markov_summary %>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(}\KeywordTok{abs}\NormalTok{(truth) <}\StringTok{ }\DecValTok{1}\NormalTok{) %>%}\StringTok{ }
\StringTok{  }\KeywordTok{group_by}\NormalTok{(simulation_type) %>%}\StringTok{ }
\StringTok{  }\KeywordTok{do}\NormalTok{(}\KeywordTok{data.frame}\NormalTok{(}
    \StringTok{`}\DataTypeTok{Type I error rate}\StringTok{`} \NormalTok{=}\StringTok{ }
\StringTok{      }\KeywordTok{error_smoother}\NormalTok{(., function(x)\{x$sig_pos |}\StringTok{ }\NormalTok{x$sig_neg\}, }\DataTypeTok{values =} \DecValTok{0}\NormalTok{)}
  \NormalTok{)) %>%}
\StringTok{  }\KeywordTok{kable}\NormalTok{(}\DataTypeTok{digits =} \DecValTok{3}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[c]{@{}lr@{}}
\toprule
simulation\_type & Type.I.error.rate\tabularnewline
\midrule
\endhead
abund & 0.258\tabularnewline
env & 0.142\tabularnewline
no\_env & 0.024\tabularnewline
\bottomrule
\end{longtable}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{pairs_summary %>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(}\KeywordTok{abs}\NormalTok{(truth) <}\StringTok{ }\DecValTok{1}\NormalTok{) %>%}\StringTok{ }
\StringTok{  }\KeywordTok{group_by}\NormalTok{(simulation_type) %>%}\StringTok{ }
\StringTok{  }\KeywordTok{do}\NormalTok{(}\KeywordTok{data.frame}\NormalTok{(}
    \StringTok{`}\DataTypeTok{Type I error rate}\StringTok{`} \NormalTok{=}\StringTok{ }
\StringTok{      }\KeywordTok{error_smoother}\NormalTok{(., function(x)\{x$sig_pos |}\StringTok{ }\NormalTok{x$sig_neg\}, }\DataTypeTok{values =} \DecValTok{0}\NormalTok{)}
  \NormalTok{)) %>%}
\StringTok{  }\KeywordTok{kable}\NormalTok{(}\DataTypeTok{digits =} \DecValTok{3}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[c]{@{}lr@{}}
\toprule
simulation\_type & Type.I.error.rate\tabularnewline
\midrule
\endhead
abund & 0.597\tabularnewline
env & 0.530\tabularnewline
no\_env & 0.308\tabularnewline
\bottomrule
\end{longtable}

@@

The top and bottom 0.5\% of the distribution have been omitted to
prevent bad behavior by the smoother in the tails. The horizontal red
line shows the nominal 95\% coverage rate and the black vertical line
marks where the ``true'' value of \(\beta\) was zero. The Type I error
rates calculated above for the Markov network closely match (one minus)
the values of these curves at zero.

@@

>>>>

{
\hypersetup{linkcolor=black}
\setcounter{tocdepth}{2}
\tableofcontents
}
\section{A: Import packages and data}\label{a-import-packages-and-data}

@@

\paragraph{Plot the regressions used above to calculate
R\^{}2}\label{plot-the-regressions-used-above-to-calculate-r2}

The plots below each correspond to one of the regressions performed
above to calculate R\^{}2 for a combination of simulation method and
estimation method. The null model (``Pairs'') was not included in these
plots because its Z-scores were frequently too large to fit on the same
scale as the other methods' estimates. Below, the thinner dashed lines
represent the 1:1 line, where the expected estimate matches the ``true''
value, and the thicker solid lines represent a linear regression through
the origin, as calculated above. In general, the four correlation- or
partial correlation-based methods had slopes that were steeper than the
1:1 line, because their estimates were constrained between -1 and 1. The
GLM and Markov network usually had slopes that were closer to 1.

(Warnings of the form ``Removed N rows containing missing values
(geom\_smooth)'' can be ignored: they are simply noting that the
regression lines were cut off in some panels of the figure).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{plot_regressions =}\StringTok{ }\NormalTok{function(type, title)\{}
  \NormalTok{x[}\KeywordTok{sample.int}\NormalTok{(}\KeywordTok{nrow}\NormalTok{(x)) , ] %>%}
\StringTok{  }\KeywordTok{subset}\NormalTok{(simulation_type ==}\StringTok{ }\NormalTok{type &}\StringTok{ }\NormalTok{method !=}\StringTok{ "null"}\NormalTok{) %>%}\StringTok{ }
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =} \NormalTok{estimate, }\DataTypeTok{y =} \NormalTok{truth)) +}
\StringTok{    }\KeywordTok{facet_wrap}\NormalTok{(~method) +}
\StringTok{    }\KeywordTok{ylim}\NormalTok{(}\KeywordTok{min}\NormalTok{(x$truth), }\KeywordTok{max}\NormalTok{(x$truth)) +}\StringTok{ }
\StringTok{    }\KeywordTok{geom_point}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{color =} \KeywordTok{factor}\NormalTok{(n_sites)), }\DataTypeTok{size =} \FloatTok{0.2}\NormalTok{) +}\StringTok{ }
\StringTok{    }\KeywordTok{geom_smooth}\NormalTok{(}\DataTypeTok{method =} \StringTok{"lm"}\NormalTok{, }\DataTypeTok{color =} \StringTok{"black"}\NormalTok{, }\DataTypeTok{formula =} \NormalTok{y ~}\StringTok{ }\DecValTok{0} \NormalTok{+}\StringTok{ }\NormalTok{x, }
                \DataTypeTok{fullrange =} \OtherTok{TRUE}\NormalTok{) +}\StringTok{ }
\StringTok{    }\KeywordTok{geom_abline}\NormalTok{(}\DataTypeTok{intercept =} \DecValTok{0}\NormalTok{, }\DataTypeTok{slope =} \DecValTok{1}\NormalTok{, }\DataTypeTok{linetype =} \DecValTok{2}\NormalTok{) +}
\StringTok{    }\KeywordTok{theme_bw}\NormalTok{() +}\StringTok{ }
\StringTok{    }\KeywordTok{ggtitle}\NormalTok{(title)}
\NormalTok{\}}

\KeywordTok{plot_regressions}\NormalTok{(}\StringTok{"no_env"}\NormalTok{, }\StringTok{"'Baseline' regressions"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning: Removed 227 rows containing missing values (geom_smooth).
\end{verbatim}

\includegraphics{show-results_files/figure-latex/unnamed-chunk-5-1.png}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot_regressions}\NormalTok{(}\StringTok{"env"}\NormalTok{, }\StringTok{"'Environment' regressions"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning: Removed 188 rows containing missing values (geom_smooth).
\end{verbatim}

\includegraphics{show-results_files/figure-latex/unnamed-chunk-5-2.png}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot_regressions}\NormalTok{(}\StringTok{"abund"}\NormalTok{, }\StringTok{"'Abundance' regressions"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning: Removed 155 rows containing missing values (geom_smooth).
\end{verbatim}

\includegraphics{show-results_files/figure-latex/unnamed-chunk-5-3.png}

@@

@@

Note that, as mentioned in the main text, inferential statistics are
only calculated for the \texttt{env} and \texttt{no\_env} simulations
(and not for the \texttt{abund} simulations).

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{pairs_summary =}\StringTok{ }\NormalTok{x[x$method ==}\StringTok{ "null"} \NormalTok{&}\StringTok{ }\NormalTok{x$simulation_type !=}\StringTok{ "abund"}\NormalTok{, ]}
\NormalTok{markov_summary =}\StringTok{ }\NormalTok{x[x$method ==}\StringTok{ "Markov network"} \NormalTok{&}\StringTok{ }\NormalTok{x$simulation_type !=}\StringTok{ "abund"}\NormalTok{, ]}

@@

The \texttt{error\_smoother} is a function that fits a kernel smoother
model to estimate the probability that a model estimate will match some
criterion (e.g. statistical significance) for some parameter as a
function of that parameter's ``true'' value. This function is used in
Figure 4C and in calculating Type I error rates below.

@@

\NormalTok{error_smoother =}\StringTok{ }\NormalTok{function(data, f, }\DataTypeTok{values =} \NormalTok{truth_seq)\{}
  \NormalTok{out =}\StringTok{ }\KeywordTok{ksmooth}\NormalTok{(}
    \DataTypeTok{x =} \NormalTok{data$truth,}
    \DataTypeTok{y =} \KeywordTok{as.numeric}\NormalTok{(}\KeywordTok{f}\NormalTok{(data)),}
    \DataTypeTok{x.points =} \NormalTok{values,}
    \DataTypeTok{kernel =} \StringTok{"normal"}\NormalTok{,}
    \DataTypeTok{bandwidth =} \FloatTok{0.25}
  \NormalTok{)}

  \NormalTok{out$y[}\KeywordTok{is.na}\NormalTok{(out$y)] =}\StringTok{ }\DecValTok{0}

  \NormalTok{out}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

@@

\CommentTok{# Calculate p(confidently wrong) for each model as a function of}
\CommentTok{# the true values}
\NormalTok{confidently_wrong =}\StringTok{ }\NormalTok{function(data)\{}
  \NormalTok{(data$sig_pos &}\StringTok{ }\NormalTok{data$truth <}\StringTok{ }\DecValTok{0}\NormalTok{) |}\StringTok{ }\NormalTok{(data$sig_neg &}\StringTok{ }\NormalTok{data$truth >}\StringTok{ }\DecValTok{0}\NormalTok{)}
\NormalTok{\}}
\NormalTok{smoothed_markov =}\StringTok{ }\KeywordTok{error_smoother}\NormalTok{(markov_summary, confidently_wrong)}
\NormalTok{smoothed_pairs =}\StringTok{ }\KeywordTok{error_smoother}\NormalTok{(pairs_summary, confidently_wrong)}

@@

\KeywordTok{plot}\NormalTok{(}
  \NormalTok{smoothed_pairs$x,}
  \NormalTok{smoothed_pairs$y,}
  \DataTypeTok{type =} \StringTok{"l"}\NormalTok{,}
  \DataTypeTok{xlab =} \StringTok{"}\CharTok{\textbackslash{}"}\StringTok{True}\CharTok{\textbackslash{}"}\StringTok{ interaction strength"}\NormalTok{,}
  \DataTypeTok{ylab =} \StringTok{"P(confidently predict wrong sign)"}\NormalTok{,}
  \DataTypeTok{bty =} \StringTok{"l"}\NormalTok{,}
  \DataTypeTok{yaxs =} \StringTok{"i"}\NormalTok{,}
  \DataTypeTok{col =} \DecValTok{2}\NormalTok{,}
  \DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, .}\DecValTok{3}\NormalTok{),}
  \DataTypeTok{lwd =} \DecValTok{2}
\NormalTok{)}
\KeywordTok{mtext}\NormalTok{(}\StringTok{"C. Error rate vs.}\CharTok{\textbackslash{}n}\StringTok{interaction strength"}\NormalTok{, }\DataTypeTok{side =} \DecValTok{3}\NormalTok{, }\DataTypeTok{adj =} \DecValTok{0}\NormalTok{, }
      \DataTypeTok{font =} \DecValTok{2}\NormalTok{, }\DataTypeTok{line =} \FloatTok{1.2}\NormalTok{, }\DataTypeTok{cex =} \NormalTok{.}\DecValTok{9}\NormalTok{) }
\KeywordTok{lines}\NormalTok{(smoothed_markov$x, smoothed_markov$y, }\DataTypeTok{lwd =} \DecValTok{2}\NormalTok{)}
\KeywordTok{legend}\NormalTok{(}\StringTok{"topleft"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{2}\NormalTok{, }\DataTypeTok{legend =} \KeywordTok{c}\NormalTok{(}\StringTok{"Null model"}\NormalTok{, }\StringTok{"Markov network"}\NormalTok{), }
       \DataTypeTok{col =} \KeywordTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{, }\DecValTok{1}\NormalTok{), }\DataTypeTok{bty =} \StringTok{"n"}\NormalTok{)}
\KeywordTok{dev.off}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

@@

\begin{verbatim}
## [1] 0.1234429
\end{verbatim}

@@

\begin{verbatim}
## [1] 0.01529933
\end{verbatim}

@@

\begin{verbatim}
## [1] 0.4640395
\end{verbatim}

@@

\begin{verbatim}
## [1] 0.22179
\end{verbatim}

@@

Type I error rates (probability of rejecting the null hypothesis given a
``true'' interaction strength of zero, according to the kernel smoother
defined by \texttt{error\_smoother} above).

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{markov_summary %>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(}\KeywordTok{abs}\NormalTok{(truth) <}\StringTok{ }\DecValTok{1}\NormalTok{) %>%}\StringTok{ }
\StringTok{  }\KeywordTok{group_by}\NormalTok{(simulation_type) %>%}\StringTok{ }
\StringTok{  }\KeywordTok{do}\NormalTok{(}\KeywordTok{data.frame}\NormalTok{(}
    \StringTok{`}\DataTypeTok{Type I error rate}\StringTok{`} \NormalTok{=}\StringTok{ }
\StringTok{      }\KeywordTok{error_smoother}\NormalTok{(., function(x)\{x$sig_pos |}\StringTok{ }\NormalTok{x$sig_neg\}, }\DataTypeTok{values =} \DecValTok{0}\NormalTok{)$y}
  \NormalTok{)) %>%}
\StringTok{  }\KeywordTok{kable}\NormalTok{(}\DataTypeTok{digits =} \DecValTok{3}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[c]{@{}lr@{}}
\toprule
simulation\_type & Type.I.error.rate\tabularnewline
\midrule
\endhead
env & 0.138\tabularnewline
no\_env & 0.024\tabularnewline
\bottomrule
\end{longtable}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{pairs_summary %>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(}\KeywordTok{abs}\NormalTok{(truth) <}\StringTok{ }\DecValTok{1}\NormalTok{) %>%}\StringTok{ }
\StringTok{  }\KeywordTok{group_by}\NormalTok{(simulation_type) %>%}\StringTok{ }
\StringTok{  }\KeywordTok{do}\NormalTok{(}\KeywordTok{data.frame}\NormalTok{(}
    \StringTok{`}\DataTypeTok{Type I error rate}\StringTok{`} \NormalTok{=}\StringTok{ }
\StringTok{      }\KeywordTok{error_smoother}\NormalTok{(., function(x)\{x$sig_pos |}\StringTok{ }\NormalTok{x$sig_neg\}, }\DataTypeTok{values =} \DecValTok{0}\NormalTok{)$y}
  \NormalTok{)) %>%}
\StringTok{  }\KeywordTok{kable}\NormalTok{(}\DataTypeTok{digits =} \DecValTok{3}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[c]{@{}lr@{}}
\toprule
simulation\_type & Type.I.error.rate\tabularnewline
\midrule
\endhead
env & 0.510\tabularnewline
no\_env & 0.298\tabularnewline
\bottomrule
\end{longtable}

@@

The top and bottom 0.5\% of the distribution have been omitted to
prevent bad behavior by the smoother (a generalized additive model) in
the tails. The horizontal red line shows the nominal 95\% coverage rate
and the black vertical line marks where the ``true'' value of \(\beta\)
was zero. The Type I error rates calculated above for the Markov network
closely match (one minus) the values of these curves at zero.

@@

\subsection{\texorpdfstring{Plot power versus true \(\beta\)
value}{Plot power versus true \textbackslash{}beta value}}\label{plot-power-versus-true-beta-value}

Calculate p(confidently detect sign of \(\beta\)) as a function of the
``true''" values. Pairs has a somewhat higher rate of rejecting the null
hypothesis in the correct direction than the Markov network, but the
benefit of this increase in power is more than offset by the large
increases in Type I errors and sign errors demonstrated above and in
Figure 4 of the main text.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{confidently_right =}\StringTok{ }\NormalTok{function(data)\{}
  \NormalTok{(data$sig_pos &}\StringTok{ }\NormalTok{data$truth >}\StringTok{ }\DecValTok{0}\NormalTok{) |}\StringTok{ }\NormalTok{(data$sig_neg &}\StringTok{ }\NormalTok{data$truth <}\StringTok{ }\DecValTok{0}\NormalTok{)}
\NormalTok{\}}

\KeywordTok{plot}\NormalTok{(}\KeywordTok{error_smoother}\NormalTok{(markov_summary, confidently_right), }
     \DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }
     \DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(-}\DecValTok{5}\NormalTok{, }\DecValTok{5}\NormalTok{),}
     \DataTypeTok{xlab =} \StringTok{"'True' coefficient value"}\NormalTok{,}
     \DataTypeTok{ylab =} \StringTok{"Power"}\NormalTok{,}
     \DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{),}
     \DataTypeTok{yaxs =} \StringTok{"i"}
\NormalTok{)}
\KeywordTok{lines}\NormalTok{(}\KeywordTok{error_smoother}\NormalTok{(pairs_summary, confidently_right), }\DataTypeTok{col =} \DecValTok{2}\NormalTok{)}
\KeywordTok{legend}\NormalTok{(}\StringTok{"topleft"}\NormalTok{, }\DataTypeTok{legend =} \KeywordTok{c}\NormalTok{(}\StringTok{"Markov network"}\NormalTok{, }\StringTok{"Pairs"}\NormalTok{), }\DataTypeTok{lty =} \DecValTok{1}\NormalTok{, }\DataTypeTok{col =} \DecValTok{1}\NormalTok{:}\DecValTok{2}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{show-results_files/figure-latex/unnamed-chunk-19-1.pdf}

### 2016-03-14 Harris_2015/simulation.Rmd
final before resubmission to Ecology?

The $alpha$ coefficients controlling species' intercepts were drawn from 
Gaussian distributions with unit variance and a mean given by `mean_alpha`.
In the R code at the end of this document, `mean_alpha` is defined as $-1$ 
for presence-absence simulations and $5$ for the abundance simulations.

The $\beta$ coefficients' magnitudes were drawn from an exponential 
distribution with rate parameter equal to 1.  The sign of each coefficient was
randomly sampled, based on the value of `p_neg`, defined below. as 0.75 (75%
negative) for the presence-absence simulations and 1.0 (100% negative) for the
simulations that included abundance.

@@

The following function calls the function above to generate "true" parameters, 
then uses them to simulate a landscape with Markov chain Monte Carlo, and 
finally saves the results to a "fakedata" folder.
I used Gibbs sampling as my Markov chain Monte Carlo transition operator.
In each round of Gibbs sampling, I cycled through all the species, randomly
updating each one's binary presence/absence vector with random samples from
a Bernoulli distribution whose conditional probability was defined by

@@

For the abundance-based simulations, each species' conditional abundance was
simulated with a Poisson distribution instead of a Bernoulli distribution. The
rate parameter of this distribution was given by 
$\mathrm{log}(1 + exp(\alpha_i + \sum_j\beta_{ij}y_j))$. (The more conventional 
exponential inverse link was not used because preliminary experiments showed that
it led to very large abundances that prevented the Gibbs sampler from mixing 
effectively.)

@@

  # Environmental states are normally distributed
  env = matrix(rnorm(n_sites * n_env), ncol = n_env)

@@

  # Landscape starts as if betas were all zero. Each species' occurrence probability
  # or abundance depends on its alpha value and on the environment (assuming alpha_env 
  # is not set to zero).
  x = matrix(
    f(rep(1, n_sites) %*% t(alpha) + env %*% alpha_env), 
    nrow = n_sites, 
    ncol = n_spp
  )

@@

Loop to simulate and save 50 landscapes for each combination of simulation type
and landscape size.

```{r}
set.seed(1)
n_spp = 20

>>>>

The $\alpha$ coefficients controlling species' intercepts were drawn from 
Gaussian distributions with unit variance and a mean given by `mean_alpha`.
In the R code at the end of this document, `mean_alpha` is defined as $-1$ 
for presence-absence simulations and $5$ for the abundance simulations.

The $\beta$ coefficients' magnitudes were drawn from an exponential 
distribution with rate parameter equal to 1.  The sign of each coefficient was
randomly sampled, based on the value of `p_neg`, defined below as 0.75 (75%
negative) for the presence-absence simulations and 1.0 (100% negative) for the
simulations that included abundance.

@@

The next function calls the function above to generate "true" parameters, 
then uses them to simulate a landscape with Markov chain Monte Carlo, and 
finally saves the results to a "fakedata" folder.
I used Gibbs sampling as my Markov chain Monte Carlo transition operator.
In each round of Gibbs sampling, I cycled through all the species,
updating each one's binary presence/absence vector with random samples from
a Bernoulli distribution whose conditional probability was defined by

@@

For the abundance-based simulations, each species' conditional abundance was
simulated with a Poisson distribution instead of a Bernoulli distribution. The
rate parameter of this distribution (which corresponds to the expected abundance)
was given by

$$y_i \sim \mathrm{log}(1 + exp(\alpha_i + \sum_j\beta_{ij}y_j))$$.

(The more conventional exponential inverse link was not used because preliminary 
experiments showed that it led to very large $y$ values, which prevented the 
Gibbs sampler from mixing effectively.)

@@

  # Environmental states are normally distributed with mean=0 and sd=1
  env = matrix(rnorm(n_sites * n_env), ncol = n_env)

@@

  # Landscape starts as if betas were all zero. Each species' occurrence 
  # probability or abundance depends on its alpha value and on the 
  # environment (assuming alpha_env is not set to zero).
  x = matrix(
    f(rep(1, n_sites) %*% t(alpha) + env %*% alpha_env), 
    nrow = n_sites, 
    ncol = n_spp
  )

@@

A loop to simulate and save 50 landscapes for each combination of simulation type
and landscape size, using the functions defined above:

```{r, eval = FALSE}
set.seed(1)
n_spp = 20
